A tutorial of Bayesian beta regressions with brms in R

### 2025-02-04 manuscript/manuscript.tex
edit case study 1

\ldots{}

@@

Here is what the tibble looks like.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ita\_egg}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 2,419 x 53
   speaker ipu    stimulus    sentence_ons sentence_off word_ons word_off v1_ons
   <chr>   <chr>  <chr>              <dbl>        <dbl>    <dbl>    <dbl>  <dbl>
 1 it01    ipu_1  Ripete 'po~         13.2         14.9     13.7     14.1   13.9
 2 it01    ipu_2  Ripete 'to~         16.9         18.6     17.4     17.9   17.5
 3 it01    ipu_3  Ripete 'pa~         20.2         21.9     20.7     21.1   20.8
 4 it01    ipu_4  Sentivo 't~         23.5         25.1     24.0     24.4   24.1
 5 it01    ipu_5  Sentivo 't~         26.3         27.8     26.8     27.2   26.9
 6 it01    ipu_6  Scrivete '~         29.2         30.9     29.7     30.1   29.8
 7 it01    ipu_7  Sentivo 'c~         32.1         33.6     32.6     33.1   32.8
 8 it01    ipu_8  Scrivete '~         35.0         36.6     35.5     35.9   35.6
 9 it01    ipu_9  Ha detto '~         41.9         43.5     42.3     42.7   42.4
10 it01    ipu_10 Sentivo 'p~         47.4         48.9     47.9     48.4   48.1
# i 2,409 more rows
# i 45 more variables: c2_ons <dbl>, v2_ons <dbl>, voice_ons <dbl>,
#   voice_off <dbl>, c1_rel <dbl>, c2_rel <dbl>, stimulus_id <dbl>,
#   sentence <chr>, word <chr>, c1 <chr>, vowel <fct>, c2 <fct>,
#   backness <chr>, height <fct>, c1_place <fct>, c2_place <fct>,
#   v1_duration <dbl>, c2_clos_duration <dbl>, rel_voff <dbl>,
#   sent_duration <dbl>, speech_rate <dbl>, speech_rate_c <dbl>, ...
\end{verbatim}

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
full summary with an explanation of each part can be found in XXX. For
each coefficient in the model, \texttt{fixef()} prints the name of the
coefficient, the mean estimate, the estimate error and the lower and
upper limits of a Bayesian Credible interval (CrI). Here, we print an
80\% CrI. There is nothing special about 95\% CrI within Bayesian
inference and instead experts recommend to check and report a variety of
CrIs. Bayesian CrIs indicate that at a certain probability levels the
``true'' estimate lies within that interval: so, for example, a 90\% CrI
{[}A, B{]} indicates that there is a 90\% probability that the ``true''
estimate is between A and B. Different probability levels correspond to
different levels of confidence: the higher the probability the higher
the confidence (always conditional on data and model). Obtaining CrIs at
different probability levels allows researchers to make more
fine-grained inferential statements than the frequentist significance
dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
in this case study but we strongly recommend researchers to always
obtain CrIs at different levels of probability and base their inferences
on all and not one in particular. To reiterate, in Bayesian inference,
an 80\% CrI indicates the range of values within which the true estimate
falls at 80\% probability or confidence. We round all values to the
nearest 2 digits for clarity.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{round}\NormalTok{(}
  \FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
  \AttributeTok{digits =} \DecValTok{2}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
              Estimate Est.Error   Q10   Q90
speech_rate_c     0.08      0.06  0.01  0.15
voweli:c2k       -0.91      0.14 -1.08 -0.74
vowele:c2k       -1.08      0.11 -1.22 -0.94
vowela:c2k       -0.99      0.12 -1.14 -0.84
vowelo:c2k       -0.79      0.14 -0.96 -0.62
vowelu:c2k       -1.00      0.16 -1.20 -0.80
voweli:c2t       -0.66      0.11 -0.79 -0.53
vowele:c2t       -0.84      0.14 -1.02 -0.66
vowela:c2t       -1.43      0.13 -1.60 -1.26
vowelo:c2t       -1.15      0.13 -1.31 -0.99
vowelu:c2t       -0.68      0.12 -0.83 -0.54
voweli:c2p       -0.68      0.11 -0.81 -0.54
vowele:c2p       -0.88      0.15 -1.07 -0.68
vowela:c2p       -1.14      0.13 -1.31 -0.98
vowelo:c2p       -0.66      0.11 -0.80 -0.53
vowelu:c2p       -0.44      0.12 -0.59 -0.28
\end{verbatim}

@@

Before being able to plot the predictions, it's important to get
familiar with the so-called posterior draws. {[}Bayesian MCMC XXX{]}.
Posterior draws can be conveniently obtained with the
\texttt{as\_draws\_df()}. For the moment, we will extract only the draws
of the constant regression coefficients (model variables starting with
\texttt{b\_}). To check which coefficients are available in a model, use
\texttt{get\_variables()} from the tidybayes package.
\texttt{as\_draws\_df()} returns a tibble where each column contains the
drawn values of a coefficient. The probability distribution of the drawn
values of each coefficient is the posterior probability distribution of
that coefficient. Note that, due to the indexing coding of vowel and C2,
all coefficient except \texttt{b\_speech\_rate\_c} are \emph{predicted
log-odds} for each vowel/C2 combination (the drawn values for
\texttt{b\_speech\_rate\_c} are drawn \emph{differences} in log-odds for
each unit increase of speech rate). The drawn values are in log-odds,
but we can convert them to proportions with \texttt{plogis()} (we will
do this when plotting below).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
\NormalTok{voi\_prop\_bm\_draws}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A draws_df: 1000 iterations, 4 chains, and 16 variables
   b_speech_rate_c b_voweli:c2k b_vowele:c2k b_vowela:c2k b_vowelo:c2k
1            0.159        -0.72        -1.05        -1.12        -1.02
2            0.035        -0.75        -1.39        -1.06        -0.89
3            0.126        -0.85        -0.98        -1.10        -0.64
4            0.064        -0.70        -1.07        -1.05        -0.83
5            0.032        -0.77        -1.18        -0.94        -0.70
6            0.073        -0.85        -1.23        -1.06        -0.77
7            0.101        -0.95        -1.07        -1.14        -0.86
8            0.074        -1.02        -1.07        -1.17        -0.83
9            0.087        -1.04        -1.07        -1.26        -0.79
10           0.092        -1.20        -1.04        -0.92        -0.64
   b_vowelu:c2k b_voweli:c2t b_vowele:c2t
1         -1.05        -0.62        -0.80
2         -1.04        -0.67        -0.99
3         -0.91        -0.66        -0.80
4         -1.07        -0.53        -0.87
5         -1.06        -0.63        -0.87
6         -1.00        -0.69        -0.73
7         -1.22        -0.74        -1.08
8         -1.22        -0.73        -1.04
9         -1.14        -0.66        -0.91
10        -0.84        -0.63        -0.61
# ... with 3990 more draws, and 8 more variables
# ... hidden reserved variables {'.chain', '.iteration', '.draw'}
\end{verbatim}

We can now wrangle this tibble and plot the posterior distributions for
each vowel/C2 combination.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 60,000 x 7
   .chain .iteration .draw type  vowel  c2     value
    <int>      <int> <int> <chr> <chr>  <chr>  <dbl>
 1      1          1     1 b     voweli c2k   -0.721
 2      1          1     1 b     vowele c2k   -1.05 
 3      1          1     1 b     vowela c2k   -1.12 
 4      1          1     1 b     vowelo c2k   -1.02 
 5      1          1     1 b     vowelu c2k   -1.05 
 6      1          1     1 b     voweli c2t   -0.625
 7      1          1     1 b     vowele c2t   -0.795
 8      1          1     1 b     vowela c2t   -1.40 
 9      1          1     1 b     vowelo c2t   -1.10 
10      1          1     1 b     vowelu c2t   -0.668
# i 59,990 more rows
\end{verbatim}

For plotting, we can use ggplot2 statistics layers from the ggdist
package. \texttt{stat\_halfeye()} plots the density of the posterior
probability (in grey), its median (point) and CrIs (lines). Let's use 60
and 80\% CrIs and transform the log-odds values to proportions with
\texttt{plogis()}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach ggdist package}
\FunctionTok{library}\NormalTok{(ggdist)}
\end{Highlighting}
\end{Shaded}

Attaching package: 'ggdist'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:brms':

    dstudent_t, pstudent_t, qstudent_t, rstudent_t
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value), vowel)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2))}
\end{Highlighting}
\end{Shaded}

@@

\caption{\label{fig-voi-prop-bm-1}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and c2 ensures that averaging applies only within draw and c2}
  \FunctionTok{group\_by}\NormalTok{(.draw, c2) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# calculate the mean value within draw/c2}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}

\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-draws-long-c2-1.pdf}}

Based on the posterior distributions of the mean voicing proportion by
consonant, /p/ has a somewhat higher voicing proportion than /k/ and
/t/. The real question is: how much higher? We can quantify this by
taking the difference of the drawn values for /p/ and those for /t, k/.
Since we want to compare /t, k/ to /p/, we should first average the
draws of /t, k/ and then take the difference of the averaged draws and
the draws of /p/.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per consonant with the mean drawn values,}
  \CommentTok{\# with one draw per raw}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ c2, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the mean of /k/ and /t/, for each draw}
    \AttributeTok{c2tk =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{c}\NormalTok{(c2k, c2t)),}
    \CommentTok{\# calculate the difference of /p/ and /t, k/}
    \AttributeTok{c2p\_tk\_diff =}\NormalTok{ c2p }\SpecialCharTok{{-}}\NormalTok{ c2tk}
\NormalTok{  )}

\NormalTok{voi\_prop\_bm\_diff}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 4,000 x 6
   .draw    c2k    c2p    c2t   c2tk c2p_tk_diff
   <int>  <dbl>  <dbl>  <dbl>  <dbl>       <dbl>
 1     1 -0.992 -0.745 -0.917 -0.953      0.208 
 2     2 -1.02  -0.790 -1.03  -0.953      0.163 
 3     3 -0.895 -0.687 -0.890 -0.953      0.266 
 4     4 -0.944 -0.814 -0.910 -0.953      0.140 
 5     5 -0.928 -0.791 -0.920 -0.953      0.162 
 6     6 -0.983 -0.728 -0.926 -0.953      0.225 
 7     7 -1.05  -0.873 -1.08  -0.953      0.0802
 8     8 -1.06  -0.875 -1.07  -0.953      0.0782
 9     9 -1.06  -0.974 -1.05  -0.953     -0.0206
10    10 -0.929 -0.829 -0.953 -0.953      0.125 
# i 3,990 more rows
\end{verbatim}

@@

\caption{\label{fig-diff-p-tk}}

@@

  \bibliography{linguistics.bib}

>>>>

The beta distribution is often considered as \emph{the} distribution for
the modeling of unit interval data
\citep{ferrari_beta_2004, cribari-neto_beta_2010}. It has been
extensively studied theoretically
\citep{krysicki_new_1999, gupta_handbook_2014, espinheira_beta_2008} and
is used a baseline to compare other unit interval distributions against
\citep{kieschnick_regression_2003, bonat_regression_2013, lopez_bayesian_2013}.
Specifically in Bayesian regression models, it was shown to perform well
across a wide range of scenarios
\citep{scholz_posterior_2023, scholz_prediction_2025}, which is why we
focus on the beta distribution in this tutorial.

We use a common mean parameterization for the beta distribution with
mean parameter \(\mu\), bounded between 0 and 1, and positive precision
parameter \(\phi\) that is roughly proportional to the inverse of the
variance \(\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}\). That is,
the larger \(\phi\) the smaller the variance of the corresponding beta
distribution. While the mathematical details are not needed to
understand the content of this tutorial, we still write down the density
below for reasons of completeness:

\[
p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
\]

where \(B\) is the Beta function, a complex integral for whose numerical
approximation efficient algorithms exist in every programming language.

@@

Table~\ref{tbl-ita-egg} shows the first ten rows of the tibble (only
relevant columns are included).

\begin{longtable}[]{@{}llllr@{}}

\caption{\label{tbl-ita-egg}First 10 rows of the data of voicing
proportion within stop closure.}

\tabularnewline

\toprule\noalign{}
speaker & word & vowel & c2 & voi\_clo\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
it01 & poto & o & t & 0.0457950 \\
it01 & topo & o & p & 0.3522513 \\
it01 & pato & a & t & 0.1440749 \\
it01 & teto & e & t & 0.3610916 \\
it01 & toto & o & t & 0.2743519 \\
it01 & puco & u & k & 0.1943984 \\
it01 & chipo & i & p & 0.2227896 \\
it01 & peco & e & k & 0.1835596 \\
it01 & poco & o & k & 0.1771007 \\
it01 & poto & o & t & 0.1525163 \\

\end{longtable}

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
full summary with an explanation of each part can be found in XXX.
Table~\ref{tbl-voi-prop-fixef} reports the output of \texttt{fixef()} as
a table (we round all values to the nearest 2 digits for clarity.). For
each coefficient in the model, \texttt{fixef()} prints the name of the
coefficient, the mean estimate, the estimate error and the lower and
upper limits of a Bayesian Credible interval (CrI). Here, we print an
80\% CrI. There is nothing special about 95\% CrI within Bayesian
inference and instead experts recommend to check and report a variety of
CrIs. Bayesian CrIs indicate that at a certain probability levels the
``true'' estimate lies within that interval: so, for example, a 90\% CrI
{[}A, B{]} indicates that there is a 90\% probability that the ``true''
estimate is between A and B. Different probability levels correspond to
different levels of confidence: the higher the probability the higher
the confidence (always conditional on data and model). Obtaining CrIs at
different probability levels allows researchers to make more
fine-grained inferential statements than the frequentist significance
dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
in this case study but we strongly recommend researchers to always
obtain CrIs at different levels of probability and base their inferences
on all and not one in particular. To reiterate, in Bayesian inference,
an 80\% CrI indicates the range of values within which the true estimate
falls at 80\% probability or confidence.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-voi-prop-fixef}Regression coefficients of a beta
regression of voicing proportion (\texttt{voi\_prop\_bm}).}

\tabularnewline

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
speech\_rate\_c & 0.08 & 0.06 & 0.01 & 0.15 \\
voweli:c2k & -0.91 & 0.14 & -1.08 & -0.74 \\
vowele:c2k & -1.08 & 0.11 & -1.22 & -0.94 \\
vowela:c2k & -0.99 & 0.12 & -1.14 & -0.84 \\
vowelo:c2k & -0.79 & 0.14 & -0.96 & -0.62 \\
vowelu:c2k & -1.00 & 0.16 & -1.20 & -0.80 \\
voweli:c2t & -0.66 & 0.11 & -0.79 & -0.53 \\
vowele:c2t & -0.84 & 0.14 & -1.02 & -0.66 \\
vowela:c2t & -1.43 & 0.13 & -1.60 & -1.26 \\
vowelo:c2t & -1.15 & 0.13 & -1.31 & -0.99 \\
vowelu:c2t & -0.68 & 0.12 & -0.83 & -0.54 \\
voweli:c2p & -0.68 & 0.11 & -0.81 & -0.54 \\
vowele:c2p & -0.88 & 0.15 & -1.07 & -0.68 \\
vowela:c2p & -1.14 & 0.13 & -1.31 & -0.98 \\
vowelo:c2p & -0.66 & 0.11 & -0.80 & -0.53 \\
vowelu:c2p & -0.44 & 0.12 & -0.59 & -0.28 \\

\end{longtable}

@@

Before being able to plot the predictions, it's important to get
familiar with the so-called posterior draws. {[}Bayesian MCMC XXX{]}.
Posterior draws can be conveniently obtained with the
\texttt{as\_draws\_df()}. For the moment, we will extract only the draws
of the constant regression coefficients (model variables starting with
\texttt{b\_}). To check which coefficients are available in a model, use
\texttt{get\_variables()} from the tidybayes package.
\texttt{as\_draws\_df()} returns a tibble where each column contains the
drawn values of a coefficient. Table~\ref{tbl-voi-prop-draws} shows the
first ten rows and first five columns of the output of
\texttt{as\_draws\_df()}. The probability distribution of the drawn
values of each coefficient is the posterior probability distribution of
that coefficient. Note that, due to the indexing coding of vowel and C2,
all coefficient except \texttt{b\_speech\_rate\_c} are \emph{predicted
log-odds} for each vowel/C2 combination (the drawn values for
\texttt{b\_speech\_rate\_c} are drawn \emph{differences} in log-odds for
each unit increase of speech rate). The drawn values are in log-odds,
but we can convert them to proportions with \texttt{plogis()} (we will
do this when plotting below).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2353}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}@{}}

\caption{\label{tbl-voi-prop-draws}First ten rows and 5 columns of the
posterior draws for the model \texttt{voi\_prop\_bm}.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
b\_speech\_rate\_c
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_voweli:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowele:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowela:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowelo:c2k
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.1593890 & -0.7213516 & -1.046577 & -1.1197185 & -1.0191525 \\
0.0351033 & -0.7492306 & -1.385450 & -1.0633164 & -0.8854320 \\
0.1259352 & -0.8472495 & -0.975776 & -1.0972735 & -0.6440836 \\
0.0640492 & -0.7033469 & -1.068626 & -1.0475190 & -0.8328252 \\
0.0319081 & -0.7659281 & -1.177656 & -0.9357499 & -0.6955130 \\
0.0729965 & -0.8494392 & -1.226692 & -1.0635364 & -0.7692243 \\
0.1009609 & -0.9546148 & -1.072955 & -1.1408255 & -0.8566750 \\
0.0735392 & -1.0160675 & -1.065025 & -1.1736263 & -0.8334884 \\
0.0870061 & -1.0436475 & -1.069337 & -1.2632061 & -0.7941857 \\
0.0923815 & -1.2042168 & -1.041238 & -0.9158689 & -0.6444492 \\

\end{longtable}

We can now wrangle this tibble and plot the posterior distributions for
each vowel/C2 combination. Table~\ref{tbl-voi-long} shows the first ten
rows.

@@

\begin{longtable}[]{@{}rrrlllr@{}}

\caption{\label{tbl-voi-long}First ten rows of posterior draws from
\texttt{voi\_prop\_bm} in long format.}

\tabularnewline

\toprule\noalign{}
.chain & .iteration & .draw & type & vowel & c2 & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voweli & c2k & -0.7213516 \\
1 & 1 & 1 & b & vowele & c2k & -1.0465766 \\
1 & 1 & 1 & b & vowela & c2k & -1.1197185 \\
1 & 1 & 1 & b & vowelo & c2k & -1.0191525 \\
1 & 1 & 1 & b & vowelu & c2k & -1.0523007 \\
1 & 1 & 1 & b & voweli & c2t & -0.6248875 \\
1 & 1 & 1 & b & vowele & c2t & -0.7951552 \\
1 & 1 & 1 & b & vowela & c2t & -1.3999773 \\
1 & 1 & 1 & b & vowelo & c2t & -1.0950180 \\
1 & 1 & 1 & b & vowelu & c2t & -0.6682030 \\

\end{longtable}

For plotting, we can use ggplot2 statistics layers from the ggdist
package. \texttt{stat\_halfeye()} plots the density of the posterior
probability (in grey), its median (point) and CrIs (lines). Let's use 60
and 80\% CrIs and transform the log-odds values to proportions with
\texttt{plogis()}. See Figure~\ref{fig-voi-prop-bm-1}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach ggdist package}
\FunctionTok{library}\NormalTok{(ggdist)}

\NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value), vowel)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2))}
\end{Highlighting}
\end{Shaded}

@@

\caption{\label{fig-voi-prop-bm-1}Expected predictions of voicing
proportion by vowel and consonant from \texttt{voi\_prop\_bm}.}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and c2 ensures that averaging applies only within draw and c2}
  \FunctionTok{group\_by}\NormalTok{(.draw, c2) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# calculate the mean value within draw/c2}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-pbdl-c2-1.pdf}}

\caption{\label{fig-voi-pbdl-c2}Expected predictions of voicing
proportion by consonant, averaged across vowels.}

Based on the expected posterior distributions of the mean voicing
proportion by consonant, /p/ has a somewhat higher voicing proportion
than /k/ and /t/. The real question is: how much higher? We can quantify
this by taking the difference of the drawn values for /p/ and those for
/t, k/ (all averaged across vowels). Since we want to compare /t, k/
with /p/, we should first average the average draws of /t, k/ and then
take the difference of the averaged /t, k/ draws and the draws of /p/.
Table~\ref{tbl-voi-pbd} shows the first ten rows of the resulting data
frame. The posterior distribution of the expected difference is shown in
Figure~\ref{fig-diff-p-tk}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per consonant with the mean drawn values,}
  \CommentTok{\# with one draw per raw}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ c2, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the mean of /k/ and /t/, for each draw}
    \AttributeTok{c2tk =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{c}\NormalTok{(c2k, c2t)),}
    \CommentTok{\# calculate the difference of /p/ and /t, k/}
    \AttributeTok{c2p\_tk\_diff =}\NormalTok{ c2p }\SpecialCharTok{{-}}\NormalTok{ c2tk}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrrrr@{}}

\caption{\label{tbl-voi-pbd}First 10 rows of expected difference of
voicing proportion between /t, k/ and /p/.}

\tabularnewline

\toprule\noalign{}
.draw & c2k & c2p & c2t & c2tk & c2p\_tk\_diff \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & -0.9918200 & -0.7450832 & -0.9166482 & -0.9533426 & 0.2082595 \\
2 & -1.0240673 & -0.7904348 & -1.0322325 & -0.9533426 & 0.1629079 \\
3 & -0.8954946 & -0.6873228 & -0.8904412 & -0.9533426 & 0.2660199 \\
4 & -0.9442361 & -0.8135191 & -0.9095122 & -0.9533426 & 0.1398235 \\
5 & -0.9278262 & -0.7908457 & -0.9200903 & -0.9533426 & 0.1624970 \\
6 & -0.9825895 & -0.7283054 & -0.9262196 & -0.9533426 & 0.2250372 \\
7 & -1.0482359 & -0.8731592 & -1.0841208 & -0.9533426 & 0.0801835 \\
8 & -1.0613713 & -0.8751219 & -1.0741555 & -0.9533426 & 0.0782208 \\
9 & -1.0622522 & -0.9739388 & -1.0481834 & -0.9533426 & -0.0205962 \\
10 & -0.9293998 & -0.8287910 & -0.9526105 & -0.9533426 & 0.1245517 \\

\end{longtable}

@@

\caption{\label{fig-diff-p-tk}Posterior distribution of the difference
between the expected voicing proportion of /t, k/ vs /p/, averaged
across vowels.}

@@

  \bibliography{linguistics.bib,statistics.bib}
